Are mental properties supervenient on brain properties?

### 2010-03-24 manuscript/supervenience.tex
incorporated jacobs comments

\tr{The primary contributions of this work are as follows.  First, a notion of supervenience emendable to empirical investigation is formally introduced.  This renders the mind-brain dualism debate a hypothesis, rather than an assumption.  Second, the limits of hypothesis testing to unravel supervenience hypotheses are provided.  Third, we posit a very general model of brains and their associated mental properties that admits a universally consistent classifier; that is, we prove that assuming the brain can be accurately represented by this model, and there is sufficient data, this algorithm is guaranteed to obtain the best possible results.  Fourth, simulations show this algorithm converges relatively rapidly.}

@@

\tr{(Note that this definition does not imply a one-to-one mapping.)}  \tr{To allow for the possibility of only \emph{partial} supervenience,} \footnote{The following was stricken from the record, as I believed it to be a non-sequitor, but i partially replace it below: Unfortunately, it is in general impossible to determine whether $L_F(g^*)=0$ without knowing $F$. Therefore,} we relax the above statistical supervenience to define $\varepsilon$-\textit{supervenience}:

@@

Practical issues regarding actually conducting the above thought experiment include: (1) as stated, we must consider the space $\mB$ to be the quotient space of graphs mod graph isomorphism, unless the vertices are {\it labeled}; (2) a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under our Frobenius norm may have a rate of convergence so slow and a computational demand so high as to be impractical; and (3) collecting enough sufficiently accurate independent and identically distributed brain-graph data and the associated mental property indicators may be beyond current technological capabilities. Regardless, related experimental work includes collecting various types of brain graph data \cite{WhiteBrenner86, DenkHorstmann04, BriggmanDenk06} and various approaches to inference on brain graphs \cite{MackeBorst08, Mishchenko09, LuLichtman09}, suggesting feasibility of such an experiment in the near future (see Appendix 2 for a simulated example of a feasible experiment). Nevertheless, our thought experiment suggests that we can hope to determine that a given mental property under consideration $\varepsilon$-supervenes on a brain's connectivity structure. \tr{This thought experiment, together with (i) the formal definition of $\varepsilon$-supervenience, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first \emph{proof} (to our knowledge) that ``hope'' of empirically investigating supervenience is possible.}

>>>>

\tr{The primary contributions of this work are as follows.  First, a notion of supervenience emendable to empirical investigation is formally introduced.  This renders the mind-brain dualism debate a hypothesis, rather than an assumption.  Second, in addition to expanding the space of questions amenable to hypothesis testing, we also demonstrate the the limits of hypothesis testing.  Third we posit a very general model of brains and their associated mental properties that admits statistical analysis in a graph theoretical and statistical framework.  Fourth, we prove that this formulation admits a universally consistent classifier (a result of general interest to graph classification, independent of the subject of the graphs).  Fifth we demonstrate through simulation that the proposed universally consistent classifer has reasonable convergence properties on simulated data.}

@@

\tr{(Note that this definition does not imply a one-to-one mapping.)}  \tr{To allow for the possibility of only \emph{partial} supervenience,} \footnote{\tr{The following was stricken from the record, as I believed it to be a non-sequitor, but i partially replace it below: Unfortunately, it is in general impossible to determine whether $L_F(g^*)=0$ without knowing $F$. Therefore,}} we relax the above statistical supervenience to define $\varepsilon$-\textit{supervenience}:

@@

Practical issues regarding actually conducting the above thought experiment include: (1) as stated, we must consider the space $\mB$ to be the quotient space of graphs mod graph isomorphism, unless the vertices are {\it labeled}; (2) a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under our Frobenius norm may have a rate of convergence so slow and a computational demand so high as to be impractical; and (3) collecting enough sufficiently accurate independent and identically distributed brain-graph data and the associated mental property indicators may be beyond current technological capabilities. Regardless, related experimental work includes collecting various types of brain graph data \cite{WhiteBrenner86, DenkHorstmann04, BriggmanDenk06} and various approaches to inference on brain graphs \cite{MackeBorst08, Mishchenko09, LuLichtman09}, suggesting feasibility of such an experiment in the near future (see Appendix 2 for a simulated example of a feasible experiment). Nevertheless, our thought experiment suggests that we can hope to determine that a given mental property under consideration $\varepsilon$-supervenes on a brain's connectivity structure. \tr{This thought experiment, together with (i) the formal definition of $\varepsilon$-supervenience, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first \emph{proof} (to our knowledge) that empirically investigating supervenience is at least theoretically possible.}
